---
title: "Import"
author: "Marius Saeltzer"
date: "24 Februar 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(openxlsx)){install.packages("openxlsx")}
if(!require(readstata13)){install.packages("readstata13")}
if(!require(knitr)){install.packages("knitr")}
if(!require(jsonlite)){install.packages("jsonlite")}
if(!require(xml2)){install.packages("xml2")}
if(!require(foreign)){install.packages("foreign")}
if(!require(rvest)){install.packages("rvest")}
```

# Organizing Data in R




# Ordering data
```{r}

c1<-read.csv("candidate_summary_2020.csv",stringsAsFactors = F)

c2<-c1[c1$Cand_Party_Affiliation=="DEM"|c1$Cand_Party_Affiliation=="REP",]
c3<-c2[c2$Cand_Office=="H",]
```

Now let's see what's inside! Are these primary candidates? Let us check the run-off structure of an example district.

https://ballotpedia.org/Alabama%27s_1st_Congressional_District_election,_2020

To get to it, we need to ORDER the dataframe:


```{r}

c3<-c3[order(c3$Cand_Office_St,c3$Cand_Office_Dist,c3$Cand_Party_Affiliation),]

names(c1)

```

As we can see, this is primary level candidate level data. But what if we are not only interested in a candidate's finances, but in a candidate's sociodemographics? What if we need data from other sources? 

## Combinining Data

The process of combining two data sets is called merging or joining data. To do so, we need a common identifier to match only things that are the same.


```{r}

library(openxlsx)
d1<-read.xlsx("Dataset Political Elites.xlsx")
d1<-d1[d1$Election=="House",]
names(d1)

```

As we can see, there are much more data points in the final campaign finance data set. It seems like the second data set mainly addresses the run-off primary candidates. Let's see whether we can combine them. 

To do so, we need a identifier variable. But this is not that easy, since d1 doesn't have a unique identifier, only names. c3 on the other hand has only a single name column. So what can we do beyond changing this by hand?

We know that the last name is split by a COMMA in the FEC data set. 

### Homogenizing a Merging Variable


It is very important to make sure the unqiue values are homogeneous and of the same type. Little differences in spelling will lead to a mismatch. The best way is to use not name, but unique id, suchn as a number stored in a string.


```{r}
names1<-strsplit(c3$Cand_Name,",")

```

This now returns a list. As we learned from dataset operations before, we can now extract this using lapply.

```{r}

c3$lastname<-unlist(lapply(names1,function(x) x[[1]]))

c3$firstname<-unlist(lapply(names1,function(x) ifelse(length(x)==1,NA,x[[2]])))


```

The second problem is that the NAME variables in d1 are not capitalized. Let's change this and create a merging variable, using toupper.

```{r}

d1$lastname<-toupper(d1$Name)
d1$firstname<-toupper(d1$First.Name)
```


Before we start, the benchmark is the number of rows in d1, as we assume that all candidates from d1 are in c3, but not all in c3 are d1.

```{r}

dc<-merge(d1,c3,by="lastname",all.y=T)

```

EWW we now have more obervations in the final dataset then the in the larger original one. This looks like a multimatch problem.

Of course, this is the case because we habe multiple surnames multiple times.


```{r}

sum(table(d1$lastname)>1)

```
We have 92 duplicate surnames.


We can also match on multiple variables!

```{r}

dc<-merge(c3,d1,by=c("lastname","firstname"))

```

Wow. 2 Matches. What now? The splitting still included some whitespaces. Let's try again.
```{r}
c3$firstname<-trimws(c3$firstname)
dc<-merge(c3,d1,by=c("lastname","firstname"))

```


This is still very bad. The first names are usually super messy as there are dozens of ways to write a first name, using different abbreviation conventions. 

To take a look at the mismatches, we can use the all argument to keep all values in the data set.
```{r}
dc<-merge(c3,d1,by=c("lastname","firstname"),all.y=T)
names(dc)

```


Now we have two strategies: homogenize the first name or choose a different variable to merge:

Little Homework: Find a way to homogenize first names! 


Let's look at another highly varying variable!


```{r}
dis<-strsplit(d1$District,"-")

d1$Cand_Office_St<-unlist(lapply(dis,function(x)x[[1]]))
d1$Cand_Office_Dist<-unlist(lapply(dis,function(x)x[[2]]))

# in c3, AT large gets number 0

d1$Cand_Office_Dist<-ifelse(d1$Cand_Office_Dist=="AL","0",d1$Cand_Office_Dist)

d1$Cand_Office_Dist<-as.numeric(d1$Cand_Office_Dist)


dc<-merge(c3,d1,by=c("lastname","Cand_Office_St","Cand_Office_Dist"))

```

Still only 745/1002


```{r}
d1$lastname<-trimws(d1$lastname)
c3$lastname<-trimws(c3$lastname)


dc<-merge(c3,d1,by=c("lastname","Cand_Office_St","Cand_Office_Dist"),all.y = T)

```

```{r}

unique(d1$Cand_Office_St)

unique(d1$Cand_Office_St)[order(unique(d1$Cand_Office_St))]
unique(c3$Cand_Office_St)[order(unique(c3$Cand_Office_St))]

unique(d1$Cand_Office_St)%in%unique(c3$Cand_Office_St)





```


Hawai was coded as HA, not HI
```{r}

d1$Cand_Office_St<-ifelse(d1$Cand_Office_St=="HA","HI",d1$Cand_Office_St)
d1$Cand_Office_St<-ifelse(d1$Cand_Office_St=="CN","CT",d1$Cand_Office_St)


dc<-merge(c3,d1,by=c("lastname","Cand_Office_St","Cand_Office_Dist"),all.y=T)

```

In contrary to our prior beliefs, the FEC database is not complete. Candidates like Jeffrey Alberts Indiana-4 are missing from the FEC, so ballotpedia seems to have more complete information!



## Level of Analysis


Data storage's most important feature is unit of analysis. Unit of analysis describes the meaning of a line or row in a data set. Unit of analysis could be the individual (in a cross sectional design), but also groups of individuals (party) or multiple observations per individual (such as panel data). 

Transforming data from a more complex file format such as xml to a rectungular form always requires decisions on the level of analysis. Are we interested in individual MPs, MPs in each election period or party aggregates. And if so, how do we store it? This part of the script will deal with these questions. 

This data set is so called panel data with individual*time as the unit of analysis. For each candidate, we have predictions over time. 

```{r}
frc<-read.csv("house_district_forecast.csv")
```

The simplest transformation in R is aggregation or summarization. It moves up the unit of analysis by loosing data. We can therefore take the mean of a time-varying variable.
In practice, is done using the formula notation: it is special way to write variable relationships in r and is used for regressions, also.


```{r}
                      # this is a tilde!      
vs<-aggregate(voteshare~candidate,frc,FUN="mean")

```  

We see that the new data set has observations which mirror the number of unique expressions of candidate. It is basically a "summary" data set. We also see that the other variables are gone. This can be solved by adding other variables WHICH VARY ON THE SAME LEVEL OF ANALYSIS.

```{r}
                      # this is a tilde!      
vs<-aggregate(voteshare~candidate+party,frc,FUN="mean")

```  

If you take another variable, which changes inside the individual over time, the computer will create an observation for each combination.

The model variable has three levels: classic, lite and deluxe. They produce different results. When we aggregate based on candidate and model, we get the average result per model per candidate.

```{r}
                      # this is a tilde!      
vs<-aggregate(voteshare~candidate+model,frc,FUN="mean")

```  

As you can see, we change the level of analysis to "larger" and loose data: all variation in predictions is flattened into a single figure. 

We could also change the level of analysis by changing where the data is stored instead of removing it. The panel format we have in the fcr data set is called a LONG data format in which each time*individual observation is a row. But you can also store the different timepoints as variables and keeping rows for individuals. This is called a WIDE format.

So if you run a simple regression on a long data format, you will run into problems with hierachical data. You can account for this by reshaping the dataset, turning differences in observations in different variables.
 

```{r}
library(reshape2)

d1<-reshape2::dcast(frc, candidate ~ model,fun=mean, value.var="voteshare",)


```


Or a little bit more useful case: let's take the polls as variables into a single data set.
```{r}
library(lubridate)

# first we round the date to make it
frc$week<-round_date(frc$forecastdate,"week")

d2<-reshape2::dcast(frc, candidate ~ week,fun=mean, value.var="voteshare",)

```

Now we can "melt" the data down into long form again.  
```{r}
d4<-reshape2::melt(d2)

```
