---
title: "Remote Access"
author: "Marius Sältzer"
date: "6 4 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Remote Access to Data Bases

As our course commences, our focus shifts from the combining of data structures to the uploading of data sets. As you all know, we will work with large datasets and will do so decentralized. This is particularly challenging in times of corona. 

To answer this challenge, I will introduce the next feature that makes SQL databases so powerful: server based remote access. For this course I got a cloud based MYSQL database from google. 

To access it, we will use the MYSQL package. It is just like SQLlite, but better suited for data transfer. 
```{r}
if(!require("RMySQL")){install.packages("RSQLite")}
library("RMySQL")

```

In contrast to storing data locally, we will now add data to an SQL server. To do so, we have to authentificate. For now, I created a number of accounts for you. You can build a connection to the data base by adding three kinds of information: 

  Host: the IP address of the Server
  user: the name of the account, your uni abbreviation
  password: preset now with your student number
  port: Channel on your computer over which to communicate
  name: Name of the database you like to access


### Whitelisted VPN Channel!

Now, servers require you to autheficate you are the one who you claim you are. I used the least annoying way of authentification, by IP address. I added all university IPs (at least I hope so) to the whitelist. Please activate your VPN clients before you try to access, or the server won't recognize you. By using the VPN client, you can enter the database through the Uni-IP network.

If you like, I can add your IP to the whitelist! Send me an email.
You can find out your IP by using this link https://www.whatismyip.com/

I need the IPv4, not the IPv6!

Let's try:

```{r}

db_user <- 'youruniid'
db_password <- 'studentIDnumber'
db_name <- 'congress'
db_host <- '35.193.111.129' # our server access
db_port <- 3306

us <-  dbConnect(RMySQL::MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)

dbListTables(us)

```


```{r}

candidata<-read.csv("alldata.csv",encoding = "UTF-8",stringsAsFactors = F)
``` 

### Primary Keys

As discussed earlier, we need to define primary keys for all tables and create the ability to join them on demand. 


1000 0001
2000 0001 


We now prepate our dataset, prepare an ID variable and clean it a tiny bit. 

```{r}

library(stringr)
      
candidata$ID<-as.numeric(paste0(1000,seq(1,nrow(candidata))))

candidata$twitter1<-gsub("https://www.twitter.com/","",candidata$twitter1)
candidata$twitter2<-gsub("https://www.twitter.com/","",candidata$twitter2)

candidata$ballotpedia.org<-NULL # the excel concat function seems to have trouble with UTF-8
names(candidata)[3]<-"First_Name"

``` 

Import the dataset completely
```{r}
dbWriteTable(us,"cand_raw",candidata,overwrite=T)

```

Now, we can create a table inside our database
```{r}
dbExecute(us,"CREATE TABLE candidates(
 ID varchar(10) primary key,
 Name TEXT,
 First_Name TEXT,
 District TEXT , 
 Party TEXT,
 Election TEXT,
 Incumbent TEXT, 
 twitter1 TEXT,
 twitter2 TEXT,
 content TEXT, 
 District_link TEXT);")

dbExecute(us,"CREATE TABLE districts(
 District varchar(5) primary key,
 District_link TEXT
          );")

```


## Inserting data

Now, as we have created the empty data sets, we can fill them with the data at hand. We can therefore migrate data from the large, unconstrained table in the more limited, structured table.

```{r}
### Check whether it worked!

d1<-dbGetQuery(us, '
SELECT *
FROM cand_raw;')


```


```{r}

dbExecute(us,"INSERT INTO candidates
 SELECT ID ,
 Name,
 First_Name,
 District, 
 Party,
 Election,
 Incumbent, 
 twitter1,
 twitter2,
 content, 
 District_link
FROM cand_raw;")
```

Now we can check if it worked!

```{r}
dbGetQuery(us, " 
           SELECT twitter1
           FROM candidates;")                

```

And remove the temporary import table

```{r}
dbExecute(us,"DROP TABLE cand_raw;")

dbDisconnect(us)

```
