---
title: "Task2"
author: "Marius Saeltzer"
date: "25 Februar 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a problem set. We start in course where your group members and me can help you if necessary.

On the homepage of the Bundeswahlleiter, we find the following file: 

# Task 1: Import

https://www.bundeswahlleiter.de/dam/jcr/72f186bb-aa56-47d3-b24c-6a46f5de22d0/btw17_kerg.csv

It looks quite messy. Find a way to import it correctly and give it meaningful column names (using names, gsub etc)

Make sure numbers and encodings are correct!

Compute the relative voteshares of each party!

Give all columns the approriate data type. 

Save it as a new csv file.
    
```{r}  

# only data
r1<-read.csv("https://www.bundeswahlleiter.de/dam/jcr/72f186bb-aa56-47d3-b24c-6a46f5de22d0/btw17_kerg.csv",sep = ";",encoding = "UTF-8",skip = 7)


# but names are wrong

r0<-read.csv("https://www.bundeswahlleiter.de/dam/jcr/72f186bb-aa56-47d3-b24c-6a46f5de22d0/btw17_kerg.csv",encoding = "UTF-8",skip = 5,sep=";",stringsAsFactors = F)

# The Problem is that the Bundeswahlleiter left a lot of blanks. Removing this requires some work

# let's first turn them into charascters
line1<-as.character(r0[1,])
line2<-as.character(r0[2,])
nline<-as.character(names(r0))

# We see that R automatically gave strange names with Xes

# Let's remove all Words that start with an X and replace by empty characters
nline<-gsub("X.*","",nline)

# now we have to fill up the "Super" catergories
# I define a nice function that checks for each element whether it is empty and replaces
# it with the PREVIOUS value (i-1)

fill_up<-function(x){
  for(i in 1:length(x)){if(x[i]==""){x[i]<-x[i-1]}}
  return(x)
}

nline<-fill_up(nline)
line1<-fill_up(line1)
line2<-fill_up(line2)

#just remove the leading NA

line1<-ifelse(line1=="NA","",line1)
line2<-ifelse(line2=="NA","",line2)


names(r1)<-paste0(nline,"_",line1,"_",line2)


## Now that we have a nice frame 

# voteshares



r2<-r1
zweitstimmen<-r2[,grepl("Zweitstimmen_Endgültig",names(r2))]

# check where voting results are stored
names(zweitstimmen) # In all but the first 4


shares<-c()
for(i in 5:ncol(zweitstimmen)){
shares<-cbind(shares,(zweitstimmen[,i]/zweitstimmen[,4])) # divide all votes through proper total votes
}

shares<-as.data.frame(shares)
names(shares)<-names(zweitstimmen)[5:ncol(zweitstimmen)]

shares<-cbind(r1[,1:5],shares)




```

## Task 2 Extract (after Session 6)

Find out which MP served the longest in Bundestag based on the Stammdaten Database.

```{r}  

list.files()

library(xml2)

r<-read_xml("MDB_STAMMDATEN.XML" )
r1<-as_list(r)

```

```{r}

# first: check how many periods each MP has in his list
  lapply(r1$DOCUMENT,function(x) length(x$WAHLPERIODEN))

  # then check what is the max value


max(unlist(lapply(r1$DOCUMENT,function(x) length(x$WAHLPERIODEN))))


# then which MP has this number 
  who<-which(unlist(lapply(r1$DOCUMENT,function(x) length(x$WAHLPERIODEN)))==
        max(unlist(lapply(r1$DOCUMENT,function(x) length(x$WAHLPERIODEN)))))


  
  unlist(lapply(r1$DOCUMENT,function(x) x$NAMEN$NAME$VORNAME))[who]
  unlist(lapply(r1$DOCUMENT,function(x) x$NAMEN$NAME$NACHNAME))[who]
  
  
  
```



Find out which candidates exited the parliament early (earlier than other members of the same period).
NOTE: THIS WAS SUPER COMPLICATED BECAUSE OF IRREGULARITIES IN THE XML. Don't worry if you didn't manage to do this.
```{r}  

# First, we need to create a dataframe that contains exits for each period 

r1$DOCUMENT[[1]]<-NULL
get_times<-function(x){
p<-as.numeric(unlist(lapply(x$WAHLPERIODEN,function(y) y$WP)))
# the authors included a null list for not ended yet. we set it to today ;)
if(19%in%p){x$WAHLPERIODEN[[which(p==19)]]$MDBWP_BIS<-"25.03.2020"}

p1<-seq(1,19) # we make a sequence vector for all of periods
exit<-unlist(lapply(x$WAHLPERIODEN,function(y) y$MDBWP_BIS))
# get all the exit dates
exit<-exit[order(p)] # sometimes wp objects are in the wrong order for some reason
p<-p[order(p)]
e1<-rep(NA,19) # create an NA vector of length 19
e1[which(p1%in%p)]<-exit # and fill in exit dates where they exist
return(e1)
}

# the result is a vector that is NA for each period the member is not in parliament 
# and an exit date if he was.

times<-lapply(r1$DOCUMENT,get_times) # we apply this to all

# transform a list into a data.frame by binding all these vectors to a matrix
tdf<-do.call(rbind,times)
tdf<-as.data.frame(tdf) # turn it into a dataframe
names(tdf)<-seq(1,19)

# now we can turn the strings into dates
for(i in 1:ncol(tdf)){
tdf[,i]<-as.Date(tdf[,i],format="%d.%m.%Y")
}

# and can compute for each time period which MPs stayed till the bitter end (value 0)

diff<-c()
for(i in 1:ncol(tdf)){
  diff<-cbind(diff,(tdf[,i]-max(tdf[,i],na.rm = T)))}


# all MPs who exited prematurely have a negative rowSum

exited<-rowSums(diff,na.rm=T)





table(exited)

exiteers<-unlist(lapply(r1$DOCUMENT,function(x) x$NAMEN$NAME$NACHNAME))[exited<0]



```


## Task 3: Merge! (after Session 6)

Merge the MP Dataset we had in class with the Election results.


Compute the individual voteshare for each MP and the share of his party's Zweitstimme (list vote) in his district. Which MPs outperform their party?
```{r}

# Bundeswahlleiter Wahlergebnisse/Election Results

bw<-read.csv("https://www.bundeswahlleiter.de/dam/jcr/72f186bb-aa56-47d3-b24c-6a46f5de22d0/btw17_kerg.csv")


# Individual level Data MP 
  # Stammdaten XML
  # Import from Bundeswahlleiter

bt<-read.csv('https://www.bundeswahlleiter.de/dam/jcr/0f2b4e8f-3057-4abf-9e43-baf593800f11/btw17_gewaehlte_utf8.csv',skip = 6,sep=";",dec = ",",encoding="UTF-8",stringsAsFactors = F)




```

