---
title: "Relational Databases II"
author: "Marius Sältzer"
date: "6 4 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About the Term Paper


Two Types:
  
  Databased Term Paper
  
    Focus on OUR data
    Own Question
    Lit Review
    Hypotheses
    Idea of Testing
  
  
  Data Paper

    Describe research idea (for BA-thesis f.e.)
    Collect your OWN data
    Build codebook
    Describe the data 
  

  Hybrid
    
    Describe a question that is not answerable with ALL our data, but relates  
    Collect extra data that is connected to OUR data  
    Add codebook and describe
    Could be a team effort, also! One substantive paper based on the new data someone else collected
    
    
## Building an SQL Database

Most of the worlds data is stored in databases and most of these databases are built in a language called SQL (Structured Query Language).

```{r}
if(!require("RSQLite")){install.packages("RSQLite")}
if(!require("dbplyr")){install.packages("dbplyr")}
```


First, we create a database. It will pop up in your working directory as a file. If it exists already, we OPEN the database as a connection. Therefore, changes in the database will immediately impact the db file as found in your folder. Adding data will make it grow.


## Relational Data bases 

We now have introduced a series of query tools, but in spite of them being stored together they have little advantage over storing data in a list of list or an XML. The real power of SQL becomes obvious when using so called keys and making a database relational. 

Next week, we will combine this data using keys, do create datasets from scratch for your RQ's! 


```{r}
install.packages("RMySQL")
library("RMySQL")

db_user <- 'student'
db_password <- 'new_password'
db_name <- 'congress'
db_table <- 'candidates'
db_host <- '127.0.0.1' # for local access
db_port <- 3306

# 3. Read data from db
mydb <-  dbConnect(RMySQL::MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)


s <- paste0("select * from ", db_table)
rs <- dbSendQuery(mydb, s)

#us <- dbConnect(RSQLite::SQLite(), "twitter2.sqlite")

```


### Adding Twitter data 

Some politicians have several twitter accounts, which means we can't use the individual key of Twitter, as we connect them to an individual candidate from 0 to 2 times. We therefore add a foreign key to the twitter data which uniquely identifies the candidate. 

```{r}

load("accounts.rdata")
load("tweets.Rdata")


dbExecute(us,"DROP TABLE twitteraccounts;")
dbExecute(us,"DROP TABLE tweets;")


accounts<-accounts[,which(sapply(accounts, class) != "list")]

accounts$text<-iconv(accounts$text,"windows-1252","UTF-8")

accounts<-accounts[,1:10]


dbWriteTable(us,"twitteraccounts",accounts,overwrite=T)





dbCreateTable(us,"tweets",all,overwrite=T)

dbListTables(us)
```


### Creating Constraints

Up until now, we created tables by importing it directly from R. However, with relational databases, we can give the columns specific meanings inside the data structure, as keys or constraints. They allow us to exactly control what happens in a dataset. 

To do so, we need to create tables with contraints

```{r}
dbExecute(us,"DROP TABLE twaccounts;")

dbExecute(us,"CREATE TABLE twaccounts(
tweet_id INTEGER,
user_id INTEGER,
screen_name TEXT,
name TEXT);")

```
Group 1:
  candidates <-unique_id
  districts  

Group 2: 
  Account
  Friends
  Tweets   status_id , user_id 
  Followers 

Group 3: 
  candidates CANDID
  raising
  spending
  committees COMMITTEEID
  

### Primary Keys


```{r}
dbExecute(us,"CREATE TABLE candidates(
 Name TEXT,
 First_Name TEXT,
 District TEXT , 
 Party TEXT,
 Election TEXT,
 Incumbent TEXT , 
 twitter1 TEXT,
 twitter2 TEXT,
 content TEXT, 
 District_link TEXT,
 PRIMARY KEY(District)
          );")

dbExecute(us,"CREATE TABLE districts(
 District TEXT ,
 District_link TEXT,
 PRIMARY KEY(District)
          );")

```

Task 1: Choose the important variables of your group's data set and create an empty table in your copy of the database. 

Send me the code.


## Inserting data

Now, as we have created the empty data sets, we can fill them with the data at hand. We can therefore migrate data from the large, unconstrained table in the more limited, structured table.

```{r}
### Import your data into the table


d1<-dbGetQuery(us, '
SELECT user_id
FROM twitteraccounts;')

names(accounts)

d1<-dbGetQuery(us, '
SELECT name
FROM personal;')



```


```{r}

dbExecute(us,"INSERT INTO twaccounts
SELECT user_id ,screen_name ,name
FROM twitteraccounts;")

dbGetQuery(us, " 
           SELECT user_id , screen_name , name
           FROM twaccounts;")                

dbExecute(us,"DROP TABLE twitteraccounts;")


```


```{r}

dbExecute(us,"DROP TABLE twittertext;")

dbExecute(us,"CREATE TABLE twittertext(
status_id INTERGER PRIMARY KEY,
user_id INTEGER,
screen_name TEXT,
text TEXT);")



dbExecute(us,"INSERT INTO twittertext
SELECT status_id, user_id ,screen_name , text 
FROM tweets;")



dbGetQuery(us, " 
           SELECT status_id, user_id ,screen_name , text
           FROM twittertext;")                


```



```{r}

```
